\newpage
\thispagestyle{empty}
\mbox{}

\newpage
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
    \textit{
    This thesis is dedicated to my mom and grandparents, \\
    I hope I have made you proud.
    }
\end{center}
\vspace*{\fill}

\newpage
\thispagestyle{empty}
\mbox{}



\chapter*{Acknowledgements}
\label{chap:acknowledgements}
\addcontentsline{toc}{chapter}{\nameref{chap:acknowledgements}}

This thesis would not have been possible without the trust, support, and companionship of my advisors, collaborators, colleagues, friends, and family.

\noindent
\textbf{Advisors.}
I would like to express my deepest gratitude to Konrad and Andreas for taking a chance on me and guiding me throughout this PhD journey. Konrad, your remarkable breadth and depth of knowledge have been truly inspirational. Your office door is always open for us, for any questions about research or life. Andreas, thank you for your trust and for allowing me to explore topics that sometimes fell outside your research interests. Your support during difficult times meant a lot to me.

I would also like to extend my gratitude to Prof. Dr. Or Litany, Prof. Dr. Angela Dai, and Prof. Dr. Simon Lucey for taking the time to review my thesis. Additionally, there are many more scientific heroes who have continually inspired me. While I can't mention everyone individually, their contributions are recognized throughout the Bibliography.

\noindent
\textbf{Collaborators.}
I feel incredibly fortunate to have worked with Zan Gojcic, who taught me what good research entails. Zan, your impressive work ethic and resourcefulness were invaluable, especially when projects were not working as expected. Collaborating with Or Litany in the later stages of my PhD was also a privilege. Thank you, Or, for hosting me during my Nvidia internship and for your remarkable insights on the NFL and DyNFL projects, you are a true scholar. Your patience and trust were essential to our success. Iro, your persistence is truly inspiring, and I hope our collaborations continue. Anton, you opened the door to the diffusion world for me. Our collaborations on Marigold and DGInStyle, as well as our engaging conversations, were a significant part of the later stage of my PhD. Jiahui, thank you for joining the PCAccumulation project. I appreciate your technical strength and all the help, especially during the final push before the deadline. Zian, thank you for inviting me to the FEGR project and teaching me everything about inverse rendering.  Rodrigo, sharing an office with you made my PhD journey brighter, thanks to our random, enriching talks. Mikhail, thanks to the bug in your code, which led to my first research paper at ETH Zurich. I miss our interesting conversations now that you're at Google. I'll definitely invite myself for Google dinner in the future. I also want to thank everyone at the Google Semantic Perception group, with whom I was fortunate to share my internship time: Federico, Keisuke, Michael Oechsle, Michael Niemeyer, Titas, and Lukas Hoyer.

To Liyuan, Yuru, Bingxin, Hanfeng, and Yuanwen, thank you for working with me on your semester projects and master's theses. Your diligence and self-motivation were truly impressive, and I wish you all a bright future on your PhD journeys.

\noindent
\textbf{Colleagues.}
I also want to extend my gratitude to all the colleagues I was fortunate to share my time with over these four years: Nico, Nikolai, Nando, Riccardo, Cenek, Torben, Alex, Binbin, Yujia, Stefano, Andrés, Ozgur, Olivier, Ghjulia, and Samantha. Thank you all for creating the best working environment and the most memorable winter retreats. Nando, let's make the presence of the Cola gangster more pronounced in the future. A special thanks goes to Monique and Jrène for their great administrative support and for organizing the winter retreats. Your support definitely made my life as a non-European person smoother. 

\noindent
\textbf{Friends.}
Ye, Jimeng, and Yikui, you have been wonderful flatmates and friends. Thank you for being supportive in so many ways. Lei Ke, Yan Wu, and the Oerlikon cats family, our Friday and weekend dinners were a source of great joy during both ups and downs. Lukas Rahmann, Leo, and Jiawen, I deeply appreciate the hiking, skiing, early-morning runs, and dinners we shared as flatmates.

Jixuan, I feel incredibly lucky to have you by my side, you make my life more colorful.