
\chapter{List of Publications}
\label{chap:pub}

\begin{enumerate}
    \item \textbf{\textbf{S. Huang}}, M. Usvyatsov, K. Schindler. ”Indoor Scene Recognition in 3D.” IROS, 2020.
    \item \textbf{S. Huang}*, Z. Gojcic*, M. Usvyatsov, A. Wieser, K. Schindler. ”PREDATOR: Registration of 3D Point Clouds with Low Overlap.” CVPR, 2021. (oral)
    \item \textbf{S. Huang}, Z. Gojcic, J. Huang, A. Wieser, K. Schindler ”Dynamic 3D Scene Analysis by Point Cloud Accumulation.” ECCV, 2022.
    \item \textbf{S. Huang}, Z. Gojcic, Z. Wang, F. William, Y. Kasten, S. Fidler, K. Schindler, O. Litany ”Neural LiDAR Fields for Novel View Synthesis.” ICCV, 2023.
    \item H. Wu, X. Zuo, S. Leutenegger, Or Litany, K. Schindle, \textbf{S. Huang} ”Dynamic LiDAR Re-simulation using Compositional Neural Fields.” CVPR, 2024. (highlight)
    \item L. Zhu, \textbf{S. Huang}, K. Schindler, I. Armeni. ”Living Scenes: Multi-object Relocalization and Reconstruction in Changing 3D Environments.” CVPR, 2024. (highlight)
    \item Z. Wang, T. Shen, J. Gao, \textbf{S. Huang}, J. Munkberg, J. Hasselgren, Z. Gojcic, W. Chen, S. Fidler ”Neural Fields meet Explicit Geometric Representations for Inverse Rendering of Urban Scenes.” CVPR, 2023.
    \item B. Ke, A. Obukhov, \textbf{S. Huang}, N. Metzger, R. Daudt, K. Schindler. ”Repurposing Diﬀusion-Based Image Generators for Monocular Depth Estimation.” CVPR, 2024. (oral, Best Paper Award Candidate)
    \item Y. Jia, L. Hoyer, \textbf{S. Huang}, T. Wang, L. Gool, K. Schindler, A. Obukhov. ”DGInStyle: DomainGeneralizable Semantic Segmentation with Image Diﬀusion Models and Stylized Semantic Control.” ECCV, 2024.
    \item L. Zhu, Y. Li, E. Sandström, \textbf{S. Huang}, K. Schindler, I. Armeni "LoopSplat: Gaussian Splatting Registration for SLAM Loop Closure." 3DV, 2025
    \item C. Stucker, B. Ke, Y. Yue, \textbf{S. Huang}, I. Armeni, K. Schindler ”ImpliCity: City Modeling from Satellite Images with Deep Implicit Occupancy Fields.” ISPRS Congress, 2022. (Best Young Author Award)
    \item L. Zhu, Y. Jia, \textbf{S. Huang}, N. Meyer, A. Wieser, K. Schindler, J. Aaron ”DeFlow: Self-supervised 3D Motion Estimation of Debris Flow.” CVPR Workshop, 2023. (Best Paper Award)
    \item T. Sun, Y. Hao, \textbf{S. Huang}, S. Savarese, K. Schindler, M. Pollefeys, I. Armeni ”Nothing stands still: A spatiotemporal benchmark on 3d point cloud registration under large geometric and temporal change.” in submission, 2024
    \item Z. Wang, \textbf{S. Huang}, J. Butt, Y. Cai, M. Varga, A. Wieser "Cross-Modal Feature Fusion for Registration of Point Clouds with Ambiguous Geometry." in submission, 2024
    \item H. Xie, A. Zhao, \textbf{S. Huang}, J. Han, S. Liu, X. Xu, X. Luo, H. Pan, Q. Du, X. Tong, "Unsupervised hyperspectral remote sensing image clustering based on adaptive density." IEEE Geoscience and Remote Sensing Letters 15 (4), 632-636
    \item H. Wu, H. Yang, \textbf{S. Huang}, D. Zeng, C. Liu, H. Zhang, C. Guo, L. Chen "Classification of point clouds for indoor components using few labeled samples." Remote Sensing 12 (14), 2181
\end{enumerate}